How Optimal is Word-Referent Identification Under Multimodal Uncertainty?

### 2019-03-15 Writeup/ms.tex
revisions

\documentclass[english,floatsintext,man]{apa6}

\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}

% Table formatting
\usepackage{longtable, booktabs}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}

\newenvironment{lltable}
  {\begin{landscape}\begin{center}\begin{ThreePartTable}}
  {\end{ThreePartTable}\end{center}\end{landscape}}

% The following enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand\getlongtablewidth{%
 \begingroup
  \ifcsname LT@\roman{LT@tables}\endcsname
  \global\longtablewidth=0pt
  \renewcommand\LT@entry[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}%
  \@nameuse{LT@\roman{LT@tables}}%
  \fi
\endgroup}

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            pdfauthor={},
            pdftitle={How Optimal is Word-Referent Identification Under Multimodal Uncertainty?},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setlength{\parindent}{0pt}
%\setlength{\parskip}{0pt plus 0pt minus 0pt}

\setlength{\emergencystretch}{3em}  % prevent overfull lines

\ifxetex
  \usepackage{polyglossia}
  \setmainlanguage{}
\else
  \usepackage[english]{babel}
\fi

% Manuscript styling
\captionsetup{font=singlespacing,justification=justified}
\usepackage{csquotes}
\usepackage{upgreek}

 % Line numbering
  \usepackage{lineno}
  \linenumbers

\usepackage{tikz} % Variable definition to generate author note

% fix for \tightlist problem in pandoc 1.14
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Essential manuscript parts
  \title{How Optimal is Word-Referent Identification Under Multimodal
Uncertainty?}

  \shorttitle{Word Identification Under Multimodal Uncertainty}

  \author{Abdellah Fourtassi\textsuperscript{1}~\& Michael C. Frank\textsuperscript{1}}

  % \def\affdep{{"", ""}}%
  % \def\affcity{{"", ""}}%

  \affiliation{
    \vspace{0.5cm}
          \textsuperscript{1} Department of Psychology, Stanford University  }

@@

    Correspondence concerning this article should be addressed to Abdellah
    Fourtassi, Postal address. E-mail:
    \href{mailto:afourtas@stanford.edu}{\nolinkurl{afourtas@stanford.edu}}
  }

  \abstract{Identifying a spoken word in a referential context requires both the
ability to integrate multimodal input and the ability to reason under
uncertainty. How do these tasks interact with one another? We introduce
a paradigm that allows us to examine how adults identify words under
joint uncertainty in the auditory and visual modalities and propose an
ideal observer model of how cues in these modalities are combined
optimally. Model predictions are tested in three experiments where word
recognition is made under two kinds of uncertainty: category ambiguity
and perceptual noise. In all cases, the optimal model explains much of
the variance in human mean judgments. When the signal is not distorted
with noise, participants weight the auditory and visual cues optimally,
that is, according to the relative reliability of each modality. But
when one modality has noise added to it, human perceivers systematically
prefer the unperturbed modality to a greater extent than the optimal
model does. The study provides a formal framework which helps to
quantify how word form and word meaning interact in word recognition
under uncertainty. Moreover it offers a first step towards a model that
accounts for form-meaning synergies in early word learning.}
  \keywords{Language understanding; audio-visual processing; word learning; speech
perception; computational modeling. \\

@@

  \usepackage[sortcites=false,sorting=none]{biblatex}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\theoremstyle{definition}
\newtheorem{example}{Example}
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}

\maketitle

\setcounter{secnumdepth}{0}

@@

Although the paradigm is within-subjects, we did not have enough
statistical power to fit a different model for each individual
participant.\footnote{We had a relatively high number of trials spanning
  all possible audio-visual matchings. Getting enough data points per
  trial per participant would have required running the online
  experimental for a much longer time (more than an hour). This could
  have increased significantly the dropout rate and possibly affected
  the quality of the data collected online.} Instead, models were
constructed with data collapsed across all participants. That being
said, we will also analyze the distribution of individual responses. The
fit was done with a nonlinear least squares regression using the NLS
package in R (Bates \& Watts, 1988). We computed the values of the
parameters, as well as their 95\% confidence intervals, through
non-parametric bootstrap (using 10000 iterations).

@@

As we noted earlier, the model addresses the question of optimality at
the population level. However, it is important to know how individual
responses are distributed. In fact, one could think of an extreme case
where optimality at the population level would be misleading. Imagine,
for instance, that in the bimodal condition half the participants relied
exclusively on the visual modality, whereas the other half relied
exclusively on the auditory modality. This case could still lead to an
aggregate behavior which appears optimal, but this optimality would be
spurious.

To examine this possibility, we consider the distribution of individual
cross-modal weighting in the bimodal condition (i.e.,
\(\frac{\sigma^2_{Vb}}{\sigma^2_{Ab}}\)). Using a factor of 10 as a
cut-off, we found that 5 participants relied almost exclusively on the
visual modality, and 12 relied almost exclusively on the auditory
modality. The percentage of both cases was relatively small compared to
the total number of participants (21.25\%). When these outliers were
removed, the distribution had a rather unimodal shape
(Figure~\ref{fig:individual}). This finding indicates that the
population's near optimality is not spurious, but based mostly on
genuine cue combination at the individual level.

As a second analysis, we asked whether the observed variance in the
individual distribution was due to mere sampling errors or whether it
corresponded to a real between-subject variability. We simulated
individual responses from the posterior distribution whose parameters
were fit to the population as a whole (i.e., the descriptive posterior).
The resulting distribution is shown in Figure~\ref{fig:individual}. For
ease of comparison, the simulated distribution was superimposed to the
real distribution. We found that the real distribution had a standard
deviation of \(sd=\) 2.24 which was larger than that of the simulated
distribution (\(sd=\) 1.14), indicating that there was real
between-subject variation beyond sampling errors. This finding means
that the participants varied in terms of how they weighted modalities:
Compared to the predictions of the population-level model, some
participants relied more on the auditory modality, whereas others relied
more on the visual modality.

In Experiment 1, we tested word recognition when there was multimodal
uncertainty in terms of category membership only. In real life, however,
tokens can undergo distortions due to noisy factors in the environment
(e.g., car noise in the background, blurry vision in a foggy weather).
In Experiment 2 and 3, we explore this additional level of uncertainty.

@@

We estimated \(b=\) -0.38 {[}-0.33, -0.42{]},
\(\sigma^2_{Ab}+\sigma^2_{Nb}=\) 9.84 {[}8.75, 11.27{]}, and
\(\sigma^2_{Vb}=\) 5.21 {[}4.84, 5.64{]}. The fit explained 0.97\% of
total variance. Cross-validation using half the responses to predict the
other half yielded \(r^2 =\) 0.95.

@@

@@

@@

@@

>>>>

\documentclass[english,,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={How Optimal is Word-Referent Identification Under Multimodal Uncertainty?},
            pdfauthor={Abdellah Fourtassi~\& Michael C. Frank},
            pdfkeywords={Language understanding; audio-visual processing; word learning; speech
perception; computational modeling.},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

  \title{How Optimal is Word-Referent Identification Under Multimodal
Uncertainty?}
    \author{Abdellah Fourtassi\textsuperscript{1}~\& Michael C.
Frank\textsuperscript{1}}
    \date{}

\shorttitle{Word Identification Under Multimodal Uncertainty}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Department of Psychology, Stanford University}
\keywords{Language understanding; audio-visual processing; word learning; speech perception; computational modeling.}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

\usepackage{longtable}
\usepackage{lscape}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage[flushleft]{threeparttable}
\usepackage{threeparttablex}

\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

\usepackage{lineno}

\linenumbers
\usepackage[sortcites=false,sorting=none]{biblatex}

@@

Correspondence concerning this article should be addressed to Abdellah
Fourtassi, Postal address. E-mail:
\href{mailto:afourtas@stanford.edu}{\nolinkurl{afourtas@stanford.edu}}}

\abstract{
Identifying a spoken word in a referential context requires both the
ability to integrate multimodal input and the ability to reason under
uncertainty. How do these tasks interact with one another? We introduce
a paradigm that allows us to examine how adults identify words under
joint uncertainty in the auditory and visual modalities and propose an
ideal observer model of how cues in these modalities are combined
optimally. Model predictions are tested in three experiments where word
recognition is made under two kinds of uncertainty: category ambiguity
and perceptual noise. In all cases, the optimal model explains much of
the variance in human mean judgments. When the signal is not distorted
with noise, participants weight the auditory and visual cues optimally,
that is, according to the relative reliability of each modality. But
when one modality has noise added to it, human perceivers systematically
prefer the unperturbed modality to a greater extent than the optimal
model does. The study provides a formal framework which helps to
quantify how word form and word meaning interact in word recognition
under uncertainty. Moreover it offers a first step towards a model that
accounts for form-meaning synergies in early word learning.

@@

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}[section]
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}
\maketitle

\begin{verbatim}
## Warning: package 'bindrcpp' was built under R version 3.4.4
\end{verbatim}

@@

Although the paradigm is within-subjects, we did not have enough
statistical power to fit a different model for each individual
participant. Instead, models were constructed with data collapsed across
all participants. The fit was done with a nonlinear least squares
regression using the NLS package in R (Bates \& Watts, 1988). We
computed the values of the parameters, as well as their 95\% confidence
intervals, through non-parametric bootstrap (using 10000 iterations).

@@

As we noted earlier, we did not have enough statistical power to fit a
different model for each individual participant. Instead, we fit one
model using the data aggregated across all participants. This analysis
is informative of behavior at the population level, only. However,
optimality at the population level could be misleading. Imagine, for
instance, that in the bimodal condition half the participants relied
exclusively on the visual modality, whereas the other half relied
exclusively on the auditory modality. This case could still lead to an
aggregate behavior which appears optimal, but this optimality would be
spurious. This, in the experiemnt below, we test the cue integration
strategy at the individual level. To this end, we increase the
statistical power by extending the length of the experiment, allowing us
to collect the number of datapoints necessary to fit the models for each
participant.

\section{Experiment 1 bis}\label{experiment-1-bis}

We recruited a planned sample of N=XX participants from Amazon
Mechanical Turk. Only participants with US IP addresses and a task
approval rate above 99\% were allowed to participate. They were paid at
an hourly rate of \$6/hour. Participants were excluded if they reported
having experienced a technical problem of any sort during the online
experiment (N=XX), or if they had less than 75\% accurate responses on
the unambiguous training trials (N=XX). The final sample consisted of N
= XX participants. {[}I should also add how many participants I excluded
because their models did not converge{]}. All participants provided
informed consent before taking the
experiment.\footnote{The sample size and exclusion criteria were specified in the pre-registration at https://osf.io/h7mzp/.}

\subsubsection{Stimuli}\label{stimuli-1}

We used the same stimuli as in Experiment 1.

\subsubsection{Design and Procedure}\label{design-and-procedure-1}

The design and procedure were similar to Experiment 1. The only
difference was that we increased the number of datapoints elicited per
subject from 70 to 300.

We fit both population-level models to the aggregated data (as in
Experiment 1), and individual models to data from each participant. For
the populaton-level models, the results (shown in
Figure~\ref{fig:unimodal}) replicated the results obtained in Experiment
1. As for the individual-level analysis, the models showed there to be
between-subject variability as seen in the paramters's distribution in
Figure XX.

\subsubsection{Bimodal condition and Cue combination
models}\label{bimodal-condition-and-cue-combination-models}

We first generated poupulation-level normative and descriptive models.
As shown in Figure~\ref{fig:bimodal}, the predicted patterns of these
models mirrored closely the results obtained in Experiment 1. In
addition, the optimality analysis (i.e., done through comparing the
optimal and descriptive models) shows that the fitted mean decision
threshold was indisguisable form the optimal threshold, thus replicating
the results of Experiment 1 (Figure~\ref{fig:bias} shows that).

At the individual level, we computed the distribution of optimal
weighing relative to the fitted (i.e., descriptive) weighing and we show
this distribution in Figure XX. The distribution centers around the
optimal prediction (i.e., 1) with a median value of XX {[}XX,XX{]}. The
variance of the distribution indicates that, while most participants
were near optimal in their cue combination, part of the participant
tended to rely predominantly on the auditory modality and another part
tended to rely predominantly on the visual modality.

{[}Here compare to random model to see whether the resuls were due to
measurement errors or to genuine betwee-subject variability, but isn't
that what the CI around the median show?{]}

This experiment was an extension to Experiment 1. Through collecting
larger-size data per subject, we were able to analyze the cue
combination optimlity, not only at the population level, but also at the
individual level. The population-level analysis replicated the results
of Experiment 1. The individual-level analysis showed that participants'
performance approximated a normal distribution centered around the
optimal combination. The large variance of this distribution indicate
that some participants tend to over-rely on the auditory modality and
others tend to over-rely on the visual modality.

In Experiment 1 and Experiment 1 bis, we tested word recognition when
there was multimodal uncertainty in terms of category membership and
inherent perceptual noise, only. In real life, however, both sound and
visual tokens can undergo distortions due to noisy factors in the
environment (e.g., car noise in the background, blurry vision in a foggy
weather). In Experiment 2 and 3, we explore this additional level of
uncertainty.

@@

We estimated \(b=\) -0.38 {[}-0.33, -0.42{]},
\(\sigma^2_{Ab}+\sigma^2_{Nb}=\) 9.84 {[}8.75, 11.27{]}, and
\(\sigma^2_{Vb}=\) 5.21 {[}4.84, 5.64{]}. The fit explained 0.97\% of
total variance. Cross-validation using half the responses to predict the
other half yielded \(r^2 =\) 0.96.

@@

\subsubsection{Participants}\label{participants-3}

@@

\subsection{Results and analysis}\label{results-and-analysis-3}

@@

\subsubsection{Unimodal conditions}\label{unimodal-conditions-4}

@@

\subsection{Discussion}\label{discussion-3}
